---
title: "ch9. Modelling Areal Daa"
author: "차서연"
date: 'Sep 6, 2020 '
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
knitr::opts_knit$set(root.dir ="C:/Users/joych/Desktop/SS/SpatialStat20/ASDA/ch9")
```


## 9.2
Areal Data : Constructing spatial weights is important
```{r}
library(spdep)
library(rgdal)
NY8 <- readOGR("C:/Users/joych/Desktop/SS/SpatialStat20/ASDA/ch9","NY8_utm18")
NY_nb <- read.gal("NY_nb.gal", region.id = row.names(NY8))
summary(NY_nb)
```

```{r}
plot(NY8, border = "grey60")
plot(NY_nb, coordinates(NY8), pch = 19, cex = 0.6, add = TRUE)
```

```{r}
Syracuse <- NY8[NY8$AREANAME == "Syracuse city",]
Sy0_nb <- subset(NY_nb, NY8$AREANAME == "Syracuse city")
summary(Sy0_nb)
```

* k nearest neighbor objects
```{r}
coords <- coordinates(Syracuse)
IDs <- row.names(Syracuse)
Sy8_nb <-knn2nb(knearneigh(coords, k=1), row.names = IDs) 
Sy9_nb <-knn2nb(knearneigh(coords, k=2), row.names = IDs) 
Sy10_nb <-knn2nb(knearneigh(coords, k=4), row.names = IDs) 
dsts <- unlist(nbdists(Sy8_nb, coords))
Sy11_nb <-dnearneigh(coords, d1 = 0, d2 = 0.75*max(dsts), row.names = IDs)
# 154, 168 no neighborhoods

```

### 9.2.2 Spatial weights objects
```{r}
Sy0_lw_W <- nb2listw(Sy0_nb)
names(Sy0_lw_W)
Sy0_lw_W$weights[2]
names(attributes(Sy0_lw_W))
```
* sums are unity

```{r}
# card: cardinality, # of neighbors
1/rev(range(card(Sy0_lw_W$neighbours)))
# 0.11111, 1.0 
summary(unlist(Sy0_lw_W$weights))
summary(sapply(Sy0_lw_W$weights, sum))
# all sums are 1
```

### binary weights

```{r}
Sy0_lw_B <- nb2listw(Sy0_nb, style = "B")
summary(unlist(Sy0_lw_B$weights))
summary(sapply(Sy0_lw_B$weights, sum))
```

### IDW weight
```{r}
dsts <- nbdists(Sy0_nb, coordinates(Syracuse))
idw <- lapply(dsts, function(x) 1/(x/1000)) # inverse 계산
Sy0_lw_idwB <- nb2listw(Sy0_nb, glist = idw, style = "B")
summary(unlist(Sy0_lw_idwB$weights))
summary(sapply(Sy0_lw_idwB$weights, sum))
```

```{r}
Sy0_lw_D1 <- nb2listw(Sy11_nb, style = "B")
Sy0_lw_D1 <- nb2listw(Sy11_nb, style = "B", zero.policy = TRUE)
print(Sy0_lw_D1, zero.policy = TRUE)
```


### Handling Spatial Weights objects
```{r}
library(Matrix)
library(spatialreg)
sum(as_dgRMatrix_listw(Sy0_lw_W)[61,])
```


### 9.2.4 Spatial autocorrelations
```{r}
set.seed(987654)
n <- length(Sy0_nb)
uncorr_x <- rnorm(n)
rho <- 0.5
autocorr_x <- invIrW(Sy0_lw_W, rho) %*% uncorr_x
```

## 9.3 Testing for Spatial Autocorrelation
```{r}
moran_u <- moran.test(uncorr_x, listw = Sy0_lw_W)
moran_a <- moran.test(autocorr_x, listw = Sy0_lw_W)
moran_a1 <- moran.test(autocorr_x, listw = nb2listw(Sy9_nb, style = "W"))
```

* NY data
```{r}
moran.test(NY8$Cases, listw = nb2listw(NY_nb))
lw_B <- nb2listw(NY_nb, style = "B")
moran.test(NY8$Cases, listw = nb2listw(NY_nb, style = "B"))
lm.morantest(lm(Cases ~ 1, NY8), listw = lw_B) # same
```
* Spatial autocorrelation exists. 

* **Saddlepoint methods**
* **Exact test**

```{r}
lm.morantest.sad(lm(Cases ~ 1, NY8), listw = lw_B)
lm.morantest.exact(lm(Cases ~ 1, NY8), listw = lw_B)

```


* **Monte Carlo** test
```{r}
set.sed(1234)
bperm <- moran.mc(NY8$Cases, listw = lw_B, nsim = 99)
```

* Poisson constant risk parametric bootstrap assessment
```{r}
r <- sum(NY8$Cases)/sum(NY8$POP8)
rni <- r * NY8$POP8
CR <- function(var, mle) rpois(length(var), lambda = mle)
MoranI.pboot <- function(var, i, listw, n, S0, ...) {
  return(moran(x = var, listw = listw, n=n, S0=S0)$I)
}
set.seed(1234)
```

```{r}
boot2 <- boot(NY8$Cases, statistic = MoranI.pboot, 
              R=999, sim = "parametric", ran.gen = CR,
              listw = lw_B, n = length(Ny8$Cases), S0 = Szero(lw_B), 

                            mle = rni)
```

```{r}
moran.plot(NY8$Cases, listw = nb2listw(NY_nb, style = "C"))
```


## 9.4 Fitting Models of Areal Data

* $Y_i$: leukaemia cases
* covariates: 3개
```{r}
nylm <- lm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8)
summary(nylm)
NY8$lmresid <- residuals(nylm)
```

* Moran test
```{r}
NYlistw <- nb2listw(NY_nb, style = "B")
lm.morantest(nylm, NYlistw)
```

```{r}
NYlistwW <- nb2listw(NY_nb, style = "W")
aple(residuals(nylm), listw = NYlistwW)
spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
         listw = NYlistwW)$lambda
```

### 9.4.1.1. SAR

```{r}
NYlistw <- nb2listw(NY_nb, style = "B")
nysar <- spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8, 
                  listw = NYlistw)
summary(nysar)
```

* weighted proportionally to the pop. size
```{r}
NY8$POP8
nylmw <- lm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8, 
                  weights = POP8)
summary(nylmw) # more significant
lm.morantest(nylmw, NYlistw)
```

```{r}
nysarw <- spatialreg::spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
                   listw = NYlistw, weights = POP8)
summary(nysarw) # AIC more lower
summary(nysar)
```

### 9.4.1.2 CAR

```{r}
nycar <- spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
                  family = "CAR", listw = NYlistw)
summary(nycar)
nycarw <- spatialreg::spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
                  family = "CAR", listw = NYlistw, weights = POP8)
summary(nycarw)
```

* Cholesky
```{r}
nysarwM <- spatialreg::spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
                  family = "SAR", listw = NYlistw, weights = POP8, 
                  method = "Matrix")
summary(nysarwM)
```

```{r}
1/range(spatialreg::eigenw(NYlistw))
nysar_11 <- spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
                     family = "SAR", listw = NYlistw, llprof = 100)
```

* SMA Model : simultaneous moving average models

```{r}
nysmaw <- spatialreg::spautolm(Z ~ NY8$PEXPOSURE + NY8$PCTAGE65P + NY8$PCTOWNHOME, data = NY8,
                   family = "SMA", listw = NYlistw, weights = POP8)
summary(nysmaw)
```

```{r}
nymix <- spatialreg::lagsarlm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME, data = NY8,
                                                            listw = NYlistw, type = "mixed")
nymix
resMCMC <- spatialreg::MCMCsamp(nysar, listw = NYlistw, mcmc = 5000, burnin = 500)
plot(resMCMC)
```

