---
title: "Ch10. Disease Mapping"
author: "차서연"
date: 'Sep 6, 2020 '
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
knitr::opts_knit$set(root.dir ="C:/Users/joych/Desktop/SS/SpatialStat20/ASDA/ch10")
```

```{r}
library(maptools)
library(spdep)
library(sf)
nc_file <- system.file("shapes/sids.shp", package="spData")[1]
llCRS <- CRS("+proj=longlat +datum=NAD27")
nc <- readShapePoly(nc_file, ID="FIPSNO", proj4string=llCRS)
rn <- sapply(slot(nc, "polygons"), function(x) slot(x, "ID"))
gal_file <- system.file("weights/ncCR85.gal", package="spData")[1]
ncCR85 <- read.gal(gal_file, region.id = rn)
```

```{r}
nc$Observed <- nc$SID74

nc$Population <- nc$BIR74

r <- sum(nc$Observed)/sum(nc$Population)
nc$Expected <- nc$Population*r


nc$SMR <- nc$Observed/nc$Expected
```


```{r}
library(RColorBrewer)
logSMR <- log(nc$SMR[nc$SMR>0])
nsteps <- 7
step <- (max(logSMR)-min(logSMR))/nsteps
brks <- exp(min(logSMR)+(0:nsteps)*step)
brks[1] <- 0
cols <- c(rev(brewer.pal(3,"Blues")), brewer.pal(4,"Reds"))
##grps
grps <- as.ordered(cut(nc$SMR, brks, include.lowest = TRUE))
plot(nc, col = cols[unclass(grps)], axes = FALSE)
box()
degAxis(1)
degAxis(2, at = c(34,35,36,37))
legend("bottomleft", legend = levels(grps), fill = cols, bty = "n", cex = 0.8, y.intersp = 0.8)
```

* $\nu, \alpha$ estimate (MME)
```{r}
library(DCluster)

eb <- empbaysmooth(nc$Observed, nc$Expected)
nc$EBPG <- eb$smthrr
eb$nu ; eb$alpha
eb$nu/eb$alpha # prior mean of the relative risk theta
```


* Probability Maps
```{r}
nc$pvalpois <- ppois(nc$Observed, nc$Expected, lower.tail = FALSE)
nbparam <- calculate.mle(as(nc, "data.frame"), model = "negbin")
nc$pvalnegbin <- pnbinom(nc$Observed, size = nbparam$size, prob = nbparam$prob, 
                         lower.tail = FALSE)
colorkeypval <- list(labels = as.character(c(0,0.01,0.05,0.1,0.5,1)), # 6 levels
                     at = (0:5)/5, height = 0.5)
pvalcols <- brewer.pal(5, "Reds")
spplot(nc, c("pvalpois", "pvalnegbin"), col.regions = rev(pvalcols), 
       at=c(0, 0.01, 0.05, 0.1, .5, 1), axes=TRUE)
```


### 10.2.2 Log-Normal Model
```{r}
ebln <- lognormalEB(nc$Observed, nc$Expected)
nc$EBLN <- exp(ebln$smthrr)

```

### 10.2.3 Global EB estimator
```{r}
EBMarshall <- EBest(nc$Observed, nc$Expected)
nc$EBMarshall <- EBMarshall[,2]
names(nc)
```


```{r}
colorkeyprob <- list(labels = levels(grps), at = (0:10)/2, height = 0.5)
cols <- c(rev(brewer.pal(3, "Blues")), brewer.pal(4, "Reds"))
spplot(nc, c("EBLN","EBMarshall", "EBPG","SMR"), col.regions = cols,
       at = brks, axes = TRUE)

atcol <- (0:5)*max(nc$SMR)/5
```

## 10.3 Spatially Structured Models
```{r}
nc$EBMrshloc <- EBlocal(nc$Observed, nc$Expected, ncCR85)$est
spplot(nc, c("EBMarshall", "EBMrshloc"), col.regions = cols, 
       at = brks, axes = TRUE)

```

## 10.4 Bayesian Hierarchical Models

```{r}
library(R2WinBUGS)
N <- nrow(nc)
d <- list(N=N, observed = nc$Observed, expected = nc$Expected)
```

```{r, eval=FALSE}
model
{
    for(i in 1:N)
    {
      observed[i] ~ dpois(mu[i])
      mu[i] <- theta[i]*expected[i]
      theta[i] ~ dgamma(nu, alpha)
    }
  
  nu ~ dgamma(0.01, 0.01)
  alpha ~ dgamma(0.01, 0.01)
}
```

### using stan
```{r}
library(rstan)
N <- nrow(nc)
d <- list(N=N, observed = nc$Observed, expected = nc$Expected)
fit1 <- stan(file = "pgmodel.stan", data = d)
```

```{r}
res <- extract(fit1)
nc$stmean <- apply(res$theta, 2, mean)
nc$stmedian <- apply(res$theta, 2, median)

```

```{r}
spplot(nc, c("SMR", "EBPG", "stmean", "stmedian"), col.regions = cols, at = brks, axes = TRUE)
```

### BayesX

```{r}
library(R2BayesX)
nc$AREAID <- 1:nrow(nc)

pgbayesx <- bayesx(Observed ~ sx(AREAID, bs = "re"),
                   offset = log(nc$Expected), family = "poisson", 
                   data = as.data.frame(nc))
```

### INLA

```{r}
library(INLA)
pginla <- inla(Observed ~ offset(log(Expected)) -1 + f(AREAID, model = "iid"), family = "poisson", 
               data = as.data.frame(nc), control.predictor = list(compute = TRUE), 
               control.compute = list(dic = TRUE))

```


### 10.4.2 Spatial Models

### CAR Model

* consider covariate : nc$nwprop
```{r}
idx <- match(attr(ncCR85.nb, "region.id"), nc$CNTY_ID)

nc.nb <- ncCR85
nc.nb <- nc.nb[order(idx)]
nc.nb <- lapply(nc.nb, function(X, idx) {
  idx[X]
}, idx = (idx))
class(nc.nb)<-"nb"
nc.nb <- nc.nb[order(idx)]
nc.nb <- nb2WB(nc.nb)

nc.nb <- nb2WB(ncCR85)
```


```{r}
nc$nwprop <- nc$NWBIR74/nc$BIR74

```

```{r}
fit1
# skewed
mcmc_areas(fit1, pars = c("theta[1]", "theta[30]"))
```

### INLA

```{r}
library(INLA)

INLA_BYM <- inla(Observed ~ nwprop + f(FIPS, model="iid") + 
  f(AREAID, model="besag", graph=nb2mat(ncCR85, style="B")) + 
  offset(log(Expected)), 
  family="poisson", data=as(nc, "data.frame"), 
  control.predictor=list(compute=TRUE))
nc$INLA <- INLA_BYM$summary.fitted.values[,1]/nc$Expected
```

### BayesX

```{r}
ncgra <- nb2gra(ncCR85)
bymbayesx <- bayesx(Observed ~ nwprop + sx(AREAID, bs = "re") + 
                      sx(FIPSNO, bs = "spatial", map = ncgra), )
```


## 10.5 Geoadditive Models

* BayesX  allows different types of smoothing Penalized spliens in the linear predictor by `sx`.
* P-splines are flexible to model non-linear effects. 

### P-spline with 10 knots
```{r}
bayessxps <- bayesx(Observed ~ sx(AREAID, bs = "ps", knots = 10),
                    offset = log(nc$Expected), family = "poisson", 
                    data = as.data.frame(nc))
```

### two-dimensional P-spline
```{r}
nc$long <- coordinates(nc)[,1]
nc$lat <- coordinates(nc)[,2]
bayesxte <- bayesx(Observed ~ sx(long, lat, bs = "te"),
                   offset = log(nc$Expected), family = "poisson", 
                   data = as.data.frame(nc))

```

